### 2020-03-04 PaperRMFI2019/Response_to_Reviewers.tex
update

\documentclass{article}
\usepackage{a4wide}
\title{Response to the Reviewer's Comments}
\author{Yansel Gonzalez Tejeda \and	Helmut Mayer}
\begin{document}

\maketitle

We thank the reviewers for identifying the weaknesses in our paper and 
providing us the opportunity to strengthen our research prior to publication.

The comments of the reviewers have been very useful, for they
made us aware that some major points we were trying to convey
remained obscure even to well-informed readers.

Following up
on the reviewers' suggestions, we have restructured the paper
by changing the order of some sections, by putting more emphasis
on motivation, and by adding clarifications at crucial spots.

In addition, we have also corrected all factual errors and
minor infelicities of presentation that the reviewers brought
to our attention. Below we indicate for each individual comment
how we have dealt with it.

\section*{Reviewer 1}
\begin{verbatim}
> 1)English and equations in the current paper are terrible. The authors need
> to improve the spelling, structure and presentation of the paper.
\end{verbatim}

\begin{verbatim}
> 2) The authors used many abbreviation words without providing any explanation.
\end{verbatim}

For all abbreviation we add the corresponding meaning. For example, Skinned 
Multi-Person Linear (SMPL),

\begin{verbatim}
> 3) There are several typo errors.
> For example:
> pg 3 :
> The SCAPE model[3] opened wide possibilities in the field. ??? which field ?
\end{verbatim}

We added the clarification: "in the field of human shape estimation".

\begin{verbatim}
> has been strongly influence --> influenced
> no human body dimensions are computed of estimated.--> from the estimated
> model
> do not require that level of detail. --> such level of detail
\end{verbatim}

 We have corrected these errors.

\begin{verbatim}
> 4) The authors used the SMPL generative human body model, developed in 2015
> to synthesize 3D human meshes.
\end{verbatim}

Yes, we do. Here we are not able to grasp the intention of the 
reviewer.

\begin{verbatim}
> 5) pg 13, it is not accurate to say "The automatic annotation process took
> approximately 1 h 47 min in an enhanced modern personal computer." What are
> the characteristics of the computer ?
\end{verbatim}

\begin{verbatim}
> 6) The description of the bloc diagram of the CNN Calvis-Net depicted in fig
> 6, page 12, is missed; The proposed CNN architecture is not convincing. No
> detail has been provided about the relation and the functioning of the
> weight, the convolutional and the FC layers.
\end{verbatim}

\begin{verbatim}
> 7) The results are far from being convincible. Experimental results are with
> limited 3D articulated meshes. It would be better if some quantified
> indicators are used (Accuracy assessment).
\end{verbatim}

We take the point, and we think our new version is much clearer.

\section*{Reviewer 2}
\begin{verbatim}
> 1). It is not clear how 8 images can be used to train CNN with ground trut
> of 3 measures.
\end{verbatim}
\begin{verbatim}
> 2). It is not clear how novel the proposed measurement calculation is.
\end{verbatim}
\begin{verbatim}
> 3). there is no comparison with existing methods
\end{verbatim}

We read the objection of the reviewer as a complaint about the
story line. Our restructuring should have remedied this.

We now emphasize the real reason why we want the reduction axioms:
as a tool for {\em compositional analysis} of the epistemic effects of
many different types of information-carrying event. It is actually
surprising that this is possible. We see this compositional analysis
as the heart of dynamic epistemic logic, not just a negotiable
'luxury' for some cases. As a side-effect, such axioms
'reduce' a complete dynamic epistemic language to its static
epistemic base -- but this is a consequence of the analysis,
not the main motivation. Of course, as we now also stress, this
reduction does have another interesting interpretation by itself:
the static language must be rich enough to 'pre-encode' the
effects of the relevenat class of informational events. This
is a requirement on optimal language design which is also known
from other areas of logic, such as conditional logic vis-a-vis
belief revision. Finally, and we may have given the wrong
impression that *this* was our main point: the reduction
analysis tends to make for simpler completeness proofs,
and some direct borrowing of known properties of the
static base into the full dynamic system.

What we also find interesting is how this methodology puts us on
the track of new epistemic operators that have not been isolated
as important before, even though they were there just below the
surface. In particular, relativized common knowledge seems a
new notion of independent interest.

\section*{Reviewer 3}
\begin{verbatim}
> Overall, the paper is very well written. However, the main contributions are
> not clear.
\end{verbatim}

\begin{verbatim}
> The segmentation method is also too constrained as it requires the
> shapes to be a pre-defined position.
\end{verbatim}

\begin{verbatim}
> The paper should also try and compare with other existing segmentation
> algorithms. This is a very well studied topic, see for example:
> https://scholar.google.com.au/scholar?
> q=3d+mesh+segmentation&hl=en&as_sdt=0&as_vis=1&oi=scholart
\end{verbatim}

While we appreciate the reviewer recommendation, it has to be said that we do 
not focus on a comparative analysis of segmentation algorithms. Nevertheless, 
we do think that we could perform such as analysis on future works.

\section*{Reviewer 4}
\begin{verbatim}
> Thank you for sharing the generated database.
\end{verbatim}

We are glad to contribute and thank the reviewer for this encouraging, 
motivational and 
constructive comment.

\section*{Reviewer 5}
\begin{verbatim}
> The main contribution of this paper is to computing the human body dimensions
> based on the geometric methods and meanwhile to testing the CNN-based
> regression method. However, I have the following concerns:
> 1. The experimental details related to the CNN-based method are missing. The
> detailed hyper-parameters used for CNN training and testing are missing.
\end{verbatim}

The reviewer is right. We add the experimental details. See also address to 
reviewer 5

\begin{verbatim}
> 2. There is no comparisons with other related methods. The error of 12mm
> achieved in this paper has no reference score.
\end{verbatim}

\begin{verbatim}
> 3. Overall, I think the problem of 3D human shape measurement is a important
> task, and the geometric method used in this paper has a certain contribution
> to this task.
\end{verbatim}

We are glad to contribute and thank the reviewer for this encouraging, 
motivational and 
constructive comment.

\begin{verbatim}
> 4. The presentation of this paper is not smooth.
\end{verbatim}

The new presentation makes our intentions clearer, we think.
The completeness proof of the logic of epistemic action models as
provided by Baltag, Moss and Solecki is complicated. Our method of
reduction axioms allows us to `lift' the elegant PDL completeness
proof to the logic of epistemic action models. And again: the
reduction axioms that we find are interesting in their own right
as we are after compositional analysis of epistemic effects of events
observed in groups of agents, and earlier dynamic epistemic logic were
simply unable to provide that methodology for common knowledge.

The `point' of LCC is to show that `really' epistemic PDL, contrary to
what the designs of intricate logics of communication and knowledge
suggest, is all that is needed for a rich logic of communication and
change. So if anything, this is a new vindication of PDL, rather than
a criticism of it.

It has to be said, though, that we do not provide a knock-down argument
why this full system is {\em needed} for our style of analysis. We have
not been able to find anything weaker that works, but our list of
open questions at the end does ask if other weaker 'solutions' exist.

\section*{Reviewer 7}
\begin{verbatim} 
> Very interesting work with a recent bibliographic study, nevertheless the
> experimentation part lacks comparative study with the existing one and
> especially in term of algorithmic complexity
\end{verbatim}

We hope that our new version is much more focussed.

The author is right on one further point. We have chosen to add
events that change the world to the standard DEL-framework, since
we feel that this increases the scope with little effort, while
still operating with the same methodology. But it is clear that
this is orthogonal to our main points. We have considered taking
it out (but that seemed a pity), or to put it into a separate
section: but that would lead to too much duplication. Our solution
now is to leave the story combined, while pointing out to the reader
who so prefers that a 'change-free' reading of our main notions and
results can be obtained by merely omitting all our technical talk
about substitutions.

\begin{verbatim}
> As far as style is concerned, the work "perspicuous" appears three
> times on the first page; once in the abstract, and twice in the
> second paragraph. While perspicuous is a delicious word, it is rare
> enough that seeing it three times makes one frown.
\end{verbatim}

We have rephrased things using a wider repertoire of adjectives.

\begin{verbatim}
> Also, typo on p.3,
> first paragraph of 2.1: "PLaza" instead of "Plaza".
\end{verbatim}

Corrected.

\begin{verbatim} 
>>REFEREE REPORT 2<<
> List of typos, corrections and comments:
>
> 1. One of the key definitions ...
\end{verbatim}

The referee is correct that this definition was wrong.

The following change solves the problem. We now take the transitive
closure rather than the *reflexive* transitive closure in the
definition. We have adapted the rest of the text including the
axioms accordingly, and now also the remark on page 5 is correct.

\end{document}